---
title: "acbs_project"
author: "Odalis Hernandez"
date: "9/22/2021"
output: html_document
---

#importing data

```{r setup, include=FALSE}
library(haven)
acbs.data <-read_spss("C:/Users/16206/Documents/MPH/Advanced Data Analysis/2018 ACBS Adult Data SPSS/ACBS_2018_ADULT_PUBLIC_LLCP.sav")
```

# Data cleaning 
```{r}
library(tidyr)
library(dplyr)
acbs.data.small<-acbs.data%>%
  select(`SEX`,`AGEG_F5_M`, `INCIDNT`,`SYMP_30D`,`COPD`, `INCOME2`,`ASTHMAGE`,`WORKENV6`, `EMP_STAT`, `@_IMPRACE`, `SMOKE100`)%>%
  zap_labels()%>%
  rename(sex= SEX,
         age=AGEG_F5_M,
         long.asth=INCIDNT,
         symp.30=SYMP_30D,
         income=INCOME2,
         work.asth=WORKENV6,
         employ=EMP_STAT,
         race=`@_IMPRACE`,
         smoke=SMOKE100)%>%
  filter(employ<2)%>%  #excluding individuals that were not at least part-time employed
  filter(sex<3)%>%
  filter(COPD<3)%>%  #excluding those who do not know COPD status
  filter(symp.30<31)%>%
  filter(income<10)%>%


  mutate(sex=recode_factor(.x=sex,
                       `1`="Male",
                       `2`="Female"))%>%
  mutate(age=recode_factor(.x=age,
                       `1`="<55",
                       `2`="<55",
                       `3`="55+",
                       `4`="55+",
                       `5`= "55+"))%>%
  mutate(long.asth=recode_factor(.x=long.asth,
                       `1`= "5 or fewer yrs ago",
                       `2`= "5 or fewer yrs ago",
                       `3`= "5+ yrs ago",
                       `7`= "Don't know",
                       `9`= NA_character_))%>%
  mutate(COPD=recode_factor(.x=COPD,
                            `1`="Yes",
                            `2`="No"))%>%
  drop_na(COPD)%>%

  mutate(income=recode_factor(.x=income,
                     `1`= "<9,999",
                     `2`= "10,000-14,999",
                     `3`="15,000-19,999",
                     `4`="20,000-24,999",
                     `5`="25,000-34,999",
                     `6`="35,000-49,999",
                     `7`="50,000-74,999",
                     `8`="75,000+"))%>%
  mutate(work.asth=recode_factor(.x=work.asth,
                       `1`= "Yes",
                       `2`="No",
                       `7`=NA_character_,
                       `9`=NA_character_))%>%
  drop_na(work.asth)%>%
  mutate(employ=recode_factor(.x=employ,
                       `1`="Employed Full-time"))%>%
  
  mutate(race=recode_factor(.x=race,
                       `1`= "White",
                       `2`= "Non-White",
                       `3`= "Non-White",
                       `4`= "Non-White",
                       `5`= "Non-White",
                       `6`= "Non-White"))%>%    #relevel race to "white" or "non-white"
  
  mutate(smoke=recode_factor(.x=smoke,
                       `1`= "Yes",
                       `2`="No",
                       `7`= NA_character_,
                       `9`=NA_character_))%>%
  mutate(ses4=case_when(income %in% c("<9,999","10,000-14,999","15,000-19,999") ~ "<19,999",
                        income %in% c("20,000-24,999","25,000-34,999","35,000-49,999") ~"20,000-49,999",
                        income %in% c("50,000-74,999") ~"50,000-74,999",
                        income %in% c("75,000+")~"75,000+"))       #releveling income to 4 levels

```


Creatubg table 1
```{r}
library(table1)

label(acbs.data.small$ses4)<-"Income"
label(acbs.data.small$age)<-"Age (under/over 55)"
label(acbs.data.small$sex)<-"Sex"
label(acbs.data.small$race)<-"Race (White or Non-White)"
label(acbs.data.small$race)<-"Race (White or Non-White)"
label(acbs.data.small$smoke)<-"Ever Smoked"
label(acbs.data.small$work.asth)<-"Asthma first caused by current job"
label(acbs.data.small$COPD)<-"COPD"

#Creation of table one showing sample frequencies
acbs.tb1<-table1(~ ses4  + sex+ smoke + age + race + work.asth | COPD, data=acbs.data.small,overall="Total", rowlabelhead="Variable")
acbs.tb1


library(openxlsx)
write.xlsx(acbs.tb1, "C:/Users/16206/Documents/MPH/Advanced Data Analysis/acbs_tb1.xlsx")   #exporting table1 to excel

```

Data visualizations 
```{r}
#bar chart displaying percent of sample with COPD stratified by exposure variable
acbs.data.small%>%
  group_by(COPD,work.asth)%>%
  count()%>%
  group_by(work.asth)%>%
  mutate(perc.copd= round(100*n/sum(n), digits = 1))%>%
  ggplot(aes(x=work.asth,y=perc.copd,fill=COPD))+
  geom_col(position="dodge", stat='identity') +
  theme_minimal()+
  ylim(0,100)+
  labs(x="Occupational Asthma",
       y="Percent",title="Asthma first caused by occupational toxins by COPD status, ACBS 2018") +
  scale_fill_manual(name = "COPD", labels = c("Has COPD", "Doesn't have COPD"), values=c("#04af57","#6d62d1"))

```

Binary logistic regression models
```{r}
#model with stepwise inclusion on biological adjustment variables
copd.RS<-glm(COPD ~ race + sex, 
                      data=acbs.data.small, 
                      family="binomial")
#lifestyle adjustment variables
copd.IS<-glm(COPD ~  ses4 + smoke, 
                      data=acbs.data.small, 
                      family="binomial")
#model with covariate adjustments only
copd.cov<-glm(COPD ~ race + ses4 + sex + smoke, 
                      data=acbs.data.small, 
                      family="binomial")
#final model with all variables and exposure
copd.all <- glm(COPD ~ race + ses4 + sex + smoke + age  + work.asth, 
                      data=acbs.data.small, 
                      family="binomial") 

library(odds.n.ends)
odds.n.ends(copd.RS)
odds.n.ends(copd.IS)
odds.n.ends(copd.cov)
odds.n.ends(copd.all)

```

Stratified logistic regression by time since asthma diagnosis, less than 5 years ago vs more than 5 years ago
```{r}
#logistic regression model- asthma diagnosis less than 5 years ago
copd.fvl <- glm(COPD ~ ses4  + sex+ smoke + age + work.asth, subset= long.asth== "5 or fewer yrs ago",
                      data=acbs.data.small, 
                      family="binomial", maxit = 100) 
#logistic regression model- asthma diagnosis more than 5 years ago
copd.fvp <- glm(COPD ~ ses4 + sex + smoke + age  + work.asth, subset= long.asth== "5+ yrs ago",
                      data=acbs.data.small,
                      family="binomial", maxit = 100) 
odds.n.ends(copd.fvl)
odds.n.ends(copd.fvp)
```

Assumptions:
For Binary Logistic regression:
*Linearity
*Influence
*Multicollinearity

All variables included in the model are categorical, so linearity and influential observations do not apply.
```{r}
#model fit
blorr::blr_model_fit_stats(copd.all)
blorr::blr_test_hosmer_lemeshow(copd.all)  #no evidence of poor model fit 

#VIF for multicolliniarity, if >2 assumption is failed
car::vif(mod = copd.all) #passed

```


